Joint training of classification model and similarity model for low-resource text classification in chatbot

### 2020-09-10 chatbot_classification/aaai_2021/paper.tex
from Zongcheng Ji.   modified:   paper.tex

  Task-specific  conversational  chatbot  \cite{wen2016network}  has been applied
  into  many practical products. For example, it can share the working burden of
  human  customer  service  agents  responsible  for  customers' questions about
  certain  products  or services. Another example is the speech-based control in
  devices  like  intelligent  speakers, such as Siri, Google home, Baidu Xiaodu,
  and smart home equipments. Regardless of whether these conversations belong to
  single-turn or multi-turn, an essential technique underlying the chatbot is to
  identify  the  real  intention  behind  a  user's  question or reply. Then the
  chatbot  is capable of conducting expected actions, text reply or accomplish some
  operations.

  It  is  a  challenging task due to some inherent properties. First,
  a customer's  utterance  within  a conversation is usually quite short. Since it
  does  not  contain  enough  information,  short  text \cite{song2014short} are
  thought  to  be  more  ambiguous  compared to long texts such as paragraphs or
  documents,  posing  a  great challenge \cite{chen2019deep} for classification
  task   \cite{phan2008learning,yan2009dynamic,hua2015short}.   Second,  in  the
  initial  stage  of  building  a  chatbot  , it is often rather hard to collect
  sufficient  labeled  data  to  train  a good model. Because developers have to
  examine tons of real system logs for typical user's utterances, and then label
  them  with expected intentions. Therefore, the key to building a task-specific
  chatbot  with  high  performance  becomes  solving the challenge of short-text
  classification   \cite{sriram2010short}   problem   under   few-shot   setting
  \cite{yu2018diverse}.

  Generally  there  are  two  kinds  of  approaches  to  address  this  problem,
  classification model that directly maps an input into a most likely label, and
  similarity  model  that searches a label whose associated data is most similar to
  an input.

@@

  The  second  step  is  \emph{negative sentence pair sampling}: As for negative
  sentence  pairs,  we will also randomly sample $P$ sentences for each class in
  the negative candidate class label set $\mathcal{C'}\backslash \{y_{i}\}$. The
  class  labels  in  $\mathcal{C'}\backslash  \{y_{i}\}$  are  the  set  of most
  confusing  class  labels comparing to the ground truth ${y_{i}}$, so we assume
  that  the  sentence  pairs  with  negative  label  grouped by user input query
  $x_{i}$  and sentences with class labels in $\mathcal{C'}\backslash \{y_{i}\}$
  are  strong  adversarial  negative  samples for sentence-pair similarity model
  that  can  help  enhance  the training speed and performance. According to the
  same  method  as  positive  sentence  pair  sampling, we can obtain the set of
  sentence  pairs  with  negative label as $\mathcal{N'}_{i}=\{[(x_{i}, x'_{1}),
  0], ..., [(x_{i}, x'_{P\cdot (K-1)}), 0]\}$

>>>>

  Task-specific  conversational  chatbot  \cite{wen2016network}  has been applied
  into  many practical products. One typical example is to share the working burden of
  human  customer  service  agents  responsible  for  customers' questions about
  certain  products  or services. Another one is the speech-based control in
  devices  like  intelligent  speakers, such as Siri, Google home, Baidu Xiaodu,
  and smart home equipments, to name a few. Regardless of whether these conversations belong to
  single or  multi-round, an essential technique underlying the chatbot is to
  identify  the  real  intention  behind  a  user's  question or reply. Then the
  chatbot  is capable of conducting expected actions, replying texts or accomplishing some
  operations.

  Identifying intention  is  a  challenging task due to some inherent properties. First,
  a customer's  utterance  within  a conversation is usually quite short. Since it
  does  not  contain  enough  information,  short  texts \cite{song2014short} are
  thought  to  be  more  ambiguous  compared to long texts such as paragraphs or
  documents,  posing  a  great challenge \cite{chen2019deep} for classification
  task   \cite{phan2008learning,yan2009dynamic,hua2015short}.   Second,  in  the
  initial  stage  of  building  a  chatbot  , it is often rather hard to collect
  sufficient  labeled  data  to  train  a good model. Because developers have to
  examine tons of real system logs for typical user's utterances, and then label
  them  with expected intentions. Therefore, to build a task-specific chatbot with
  high  performance, it is essential to solve the challenge of short-text
  classification   \cite{sriram2010short}   problem   under   few-shot   setting
  \cite{yu2018diverse}.

  Generally  there  are  two  kinds  of  approaches  to  address  this  problem,
  one is the classification model that directly maps an input into a most likely label, and
  the other one is the similarity  model  that searches a label whose associated data is most similar to
  an input.

@@

  The  second  step  is  \emph{negative sentence pair sampling}: As for negative
  sentence  pairs,  we will also randomly sample $P$ sentences for each class in
  the negative candidate class label set $\mathcal{C'}\backslash \{y_{i}\}$. The
  class  labels  in  $\mathcal{C'}\backslash  \{y_{i}\}$  are  the  set  of most
  confusing  class  labels comparing to the ground truth ${y_{i}}$, so we assume
  that  the  sentence  pairs  with  negative  label  grouped by user input query
  $x_{i}$  and sentences with class labels in $\mathcal{C'}\backslash \{y_{i}\}$
  are  strong  adversarial  negative  samples for sentence-pair similarity model
  that  can  help  enhance  the training speed and performance. According to the
  same  method  as  positive  sentence  pair  sampling, we can obtain the set of
  sentence  pairs  with  negative label as $\mathcal{N'}_{i}=\{[(x_{i}, x'_{1}),
  0], ..., [(x_{i}, x'_{P\cdot K}), 0]\}$
